Something about Performance, 2D Component of NEMO. (TBC)

### 2016-07-08 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolited2d paper: merge in updates. Roofline plot is now for E5-1620 rather than Archer.

\author{J.~Appleyard, M.~Ashworth, R.~W.~Ford, J.~Holt, H.~Liu, A.~R.~Porter and G.~D.~Riley% <-this % stops a space
\thanks{J.~Appleyard is with NVIDIA.}%
\thanks{H.~Liu and J.~Holt are with the National Oceanography Centre, Liverpool, UK.}%
\thanks{M.~Ashworth, R.~W.~Ford and A.~R.~Porter are with the Science and Technology Facilities Council, UK.}% <-this % stops a space
\thanks{G.~Riley is with the University of Manchester, Manchester, UK.}% <-this % stops a space
}

@@

Once the problem size is increased to $128^2$, a single time-step
takes roughly 200~$\mu$s and only the Intel-compiled OpenMP version is
comparable in performance to the OpenACC version. For all of the
larger problem sizes plotted in Figure~\ref{FIG_cpu_cf_gpu} the GPU
version is considerably faster than the CPU.  For problem sizes of
$1024^2$ and greater, the 30MB cache of the Ivy Bridge CPU is
exhausted and performance becomes limited by the bandwidth to main
memory. At this stage the OpenACC version of the code on the GPU is
some {\bf 3.7 (this is the factor for the CUDA version!)}sssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssssss times faster
than the OpenMP version on the CPU. {\bf Can we relate this to the
  difference in peak memory b/w between the GPU and CPU?}

@@

Although we have investigated how the performance of the \psykal
version of nemolite2d compares with that or the original, we have not
addressed how efficient the original actually is. In order to do so we
use the Roofline Model~\cite{roofline} which provides a relatively
simple way of characterising the performance of a code in terms of
whether it is memory-bandwidth bound or compute bound. To do so we
follow the approach suggested in~\cite{para_pearls} and use the
well-known STREAM~\cite{stream} and LINPACK~\cite{linpack} benchmarks
in order to get the upper bounds on the memory bandwidth and FLOPS for
the test hardware. Since we are using an Intel Ivybridge CPU we used
the Intel Math Kernel Library implementation of LINPACK.

@@

\begin{figure}
\centering
\includegraphics[width=\picwidth]{roofline}
\caption{Comparison of the performance achieved by nemolite2d and Shallow}
\label{FIG_roofline}
\end{figure}

>>>>

\author{A.~R.~Porter, J.~Appleyard, M.~Ashworth, R.~W.~Ford, J.~Holt, H.~Liu and G.~D.~Riley% <-this % stops a space
\thanks{M.~Ashworth, R.~W.~Ford and A.~R.~Porter are with the Science and Technology Facilities Council, UK.}% <-this % stops a space
\thanks{J.~Appleyard is with NVIDIA.}%
\thanks{H.~Liu and J.~Holt are with the National Oceanography Centre, Liverpool, UK.}%
\thanks{G.~Riley is with the University of Manchester, Manchester, UK.}% <-this % stops a space
}

@@

Once the problem size is increased to $128^2$, a single time-step
takes roughly 200~$\mu$s and only the Intel-compiled OpenMP version is
comparable in performance to the OpenACC version. For all of the
larger problem sizes plotted in Figure~\ref{FIG_cpu_cf_gpu} the GPU
version is considerably faster than the CPU.  For problem sizes of
$1024^2$ and greater, the 30MB cache of the Ivy Bridge CPU is
exhausted and performance becomes limited by the bandwidth to main
memory. At this stage the OpenACC version of the code on the GPU is
some {\bf 3.7 (this is the factor for the CUDA version!)} times faster
than the OpenMP version on the CPU. {\bf Can we relate this to the
  difference in peak memory b/w between the GPU and CPU?}

@@

Although we have investigated how the performance of the \psykal
version of nemolite2d compares with that or the original, we have not
addressed how efficient the original actually is. In order to do so we
use the Roofline Model~\cite{roofline} which provides a relatively
simple way of characterising the performance of a code in terms of
whether it is memory-bandwidth bound or compute bound. To do so we
follow the approach suggested in~\cite{para_pearls} and use the
 STREAM~\cite{stream} and LINPACK~\cite{linpack} benchmarks
in order to get the upper bounds on the memory bandwidth and FLOPS for
the test hardware. Since we are using an Intel Ivybridge CPU we used
the Intel Math Kernel Library implementation of LINPACK.

@@

\begin{figure}
\centering
\includegraphics[width=\picwidth]{roofline}
\caption{Comparison of the performance achieved by NEMOLite2D and
  Shallow on the E5-1620 CPU. Results for the former are for the
  $256^2$ domain since that gave the best performance.}
\label{FIG_roofline}
\end{figure}

In Figure~\ref{FIG_roofline} we show the performance of kernels from
both Shallow and NEMOLite2D on the roofline model for the E5-2697 CPU.
This demonstrates that the Shallow kernel is achieving a performance
roughly consistent with saturating the available memory bandwidth. In
contrast, the kernel taken from NEMOLite2D is struggling to reach a
performance consistent with saturating the bandwidth to main
memory. The fact that reducing the problem size does not improve the
performance points to more fundamental issues with the way that the
kernel is implemented.

We investigated this using the simpler Continuity kernel since it too
performs poorly while being much shorter.
{\bf Plot of Cont. GFLOPS as function of optimisations here.}
\begin{itemize}
\item AVX versus SSE versus no SIMD vectorisation.
\item Telling the compiler about alignment
\item The division operation
\end{itemize}
Move back to Momentum kernel. Discuss further issues:
\begin{itemize}
\item IFs within loop (masking)
\end{itemize}
